% !TeX root = ../xdupgthesis_template_lc.tex

Roadside sensing technology provides real-time access to traffic operating states from an infrastructure-wide perspective and serves as a key support for traffic organization optimization, refined parking management, and intelligent transportation systems. As urban traffic becomes increasingly complex and traffic governance becomes more fine-grained, the demand for low-cost and highly reliable roadside vehicle sensing data that supports long-term online operation is becoming more urgent. Vehicle-type information, together with parking and road occupancy states, provides an important basis for traffic performance evaluation, parking management, and road-space governance. Compared with video- and radar-based approaches, geomagnetic sensors offer advantages such as low cost, ease of deployment, and all-weather operation, making them suitable for long-term roadside monitoring. However, a single geomagnetic node provides only limited information and is vulnerable to slow environmental magnetic drift, variations in vehicle speed, continuous traffic disturbances, and pseudo-stable states, which make feature extraction and state recognition challenging. To address these issues, this thesis investigates vehicle classification and parking and occupancy state determination using continuous 50\,Hz observations from a roadside single-point triaxial geomagnetic sensor under a unified event-level input framework. The main contributions are summarized as follows.

First, to address the difficulty of vehicle recognition caused by vehicle-speed variation and waveform distortion, a three-class vehicle classification method that balances lightweight edge deployment with high offline accuracy is proposed. A geomagnetic vehicle-event dataset collected from real traffic scenes is first constructed. For edge-side deployment, multidimensional statistical features reflecting magnetic disturbance amplitude, energy, temporal area, and morphological distribution are extracted and fused to build a low-overhead Softmax-based baseline classifier. For offline enhancement, dynamic time warping is introduced to align temporal features under speed-induced waveform stretching and local misalignment, and a one-dimensional convolutional neural network is designed for deep feature extraction and enhanced classification. Experimental results show that the proposed framework satisfies the computational constraints of edge devices while effectively improving class separability and mitigating confusion between adjacent vehicle categories.

Second, for parking and occupancy recognition in road scenarios, a dynamic-reference decision method based on signal stable points is proposed. To address the failure of fixed references caused by slow environmental drift, the difficulty of obtaining stable evidence in time under continuous traffic, and the tendency of pseudo-stable states to cause boundary misjudgment, a parking and occupancy detection method based on stable-point extraction and dynamic-reference maintenance is designed. On this basis, a similarity-gating mechanism, an outer finite-state-machine architecture, and a degradation branch for continuous traffic are incorporated to enable unified determination of parking confirmation, occupancy maintenance, and release boundaries. Experimental results show that the proposed method achieves good detection accuracy under normal conditions and retains strong robustness and resistance to interference under challenging conditions such as dense continuous traffic and geomagnetic slow drift.

Overall, this thesis focuses on roadside vehicle sensing with a single geomagnetic sensor and investigates a three-class vehicle classification method together with a road parking and occupancy detection method supported by vehicle event detection. The results indicate that, under the constraints of low cost, low power consumption, and ease of deployment, a single geomagnetic sensor has strong potential to support roadside vehicle recognition and parking-occupancy-state sensing, and can provide a reliable theoretical basis and methodological reference for the refined governance of smart roads.
